---
output:
  pdf_document: default
  html_document: default
---

---
title: "Report MovieLens project"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))

movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data

set.seed(1, sample.kind="Rounding")
# if using R 3.5 or earlier, use `set.seed(1)` instead
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set

validation <- temp %>% 
     semi_join(edx, by = "movieId") %>%
     semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set

removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
validation_New <- validation  
validation <- validation %>% select(-rating)


```

##Introduction 

In this project we use dataset Movielens included in the dslabs package.
The Movilens is composed from users (userId) that give a  rating between 0 and 5 (rating) in a specific date and time (timestamp) for the movies (movieId) that have a title and a genres associated.
Our goal is using the inputs in one subset to predict movie ratings in the validation set that will compared with RMSE.

## Data exploration

The first step, is to see the structur of our data (training dataset)


```{r edx1}
head(edx)
```

the data set edx, is composed by 6 six variable, of two type:
1) Quantitave variable:userId (number identify the user), movieId (number identify the movie), timestamp (number that iden identify date and time), rating (valutation of ranting movies - that is a discete variable, that hava a value from 0.5 to 5)
2) Qualitative variable: title (name movie title - not unique), genres (type genres associated with the movie).

```{r edx2}
str(edx)
```

this is summaries of every variable

```{r edx3}
summary(edx)
```


## Count votes of genre

We see the top ten genres that have more review

```{r pressure, echo=FALSE}
top_10_genre<- edx%>%
  group_by(genres) %>%
  summarize(count = n()) %>%
  arrange(desc(count))  %>% top_n(10)  

top_10_genre

```



## Count vote for user

This is a histogram that rapresent the number votes give for every user.

```{r pressure2, echo=FALSE}
edx %>% count(userId) %>% 
  ggplot(aes(n)) + 
  geom_histogram(bins = 40, color = "red",fill="black")+
labs(title="Count vote for user")
```


## Top 10 movies with most vote
Here we have the list of top ten movie with more movie review
 
```{r pressure3, echo=FALSE}
top_10_film<- edx %>% group_by(title) %>% summarize(count=n()) %>% arrange(desc(count)) %>% top_n(10)  
top_10_film
```


## Model prediction
We applicate different model to predict the rating movies and we select one that have more lower RMSE (Residual Mean Standard Error)

- In a first model we use the mean of rating for predict the rating of movies
This model find the mean of training set of reating movies

```{r edx4}
mu_edx <- mean(edx$rating)  
mu_edx

rmse_results <- data_frame(method = "Only Mean", RMSE = mu_edx)

```


and the quality of model is:

```{r edx5}
basic_rmse <- RMSE(validation_New$rating,mu_edx)
basic_rmse
```



- In a second model we applicate the penalty of the movie effect

```{r edx6}
moviePenalty <- edx %>% 
  group_by(movieId) %>% 
  summarize(b_movie = mean(rating - mu_edx))

moviePenalty

```

and we have the quality of predict model is

```{r edx7}
predict_ratings_movie<- validation %>% 
  left_join(moviePenalty, by='movieId') %>%
  mutate(pred = mu_edx + b_movie) 
modelMovies_rmse <- RMSE(validation_New$rating,predict_ratings_movie$pred)
rmse_results <- bind_rows(rmse_results, data_frame(method="Movie Effect Model", RMSE = modelMovies_rmse ))

modelMovies_rmse
```


``````{r moviePenalaty, echo=FALSE}

moviePenalty %>% qplot(b_movie, geom ="histogram", bins = 40, data = ., color = "black", main = "Penalty Movie")
moviePenalty
```


- In the third model, for predict tha rating we use the penalty of movie effect (the previous model) and the penalty of users effect 

Before we calculate the penalty of users
```{r PenaltyUser}
penaltyUser <- edx %>% 
  left_join(moviePenalty, by='movieId') %>%
  group_by(userId) %>%
  summarize(b_user = mean(rating - mu_edx - b_movie))

penaltyUser



```



and now we can calculate the RMSE of this model, that is

```{r edx9}


predicted_ratings_user <- validation %>% 
  left_join(moviePenalty, by='movieId') %>%
  left_join(penaltyUser, by='userId') %>%
  mutate(pred = mu_edx + b_movie + b_user) 
# test rmse results 
model_MoviesUsers_rmse <- RMSE(validation_New$rating,predicted_ratings_user$pred)
rmse_results <- bind_rows(rmse_results,data_frame(method="Movie and User Effect Model", RMSE = model_MoviesUsers_rmse ))
model_MoviesUsers_rmse
```

- The fourth model, we consider users,movies,years, genres and than applicate the regularization 
```{r pressure4, echo=FALSE}
lambdas <- seq(0, 20, 0.25)

rmses <- sapply(lambdas, function(l){
  
  mu_edx <- mean(edx$rating)
  
  b_movie <- edx %>% 
    group_by(movieId) %>%
    summarize(b_movie = sum(rating - mu_edx)/(n()+l))
  
  b_user <- edx %>% 
    left_join(b_movie, by="movieId") %>%
    group_by(userId) %>%
    summarize(b_user = sum(rating - b_movie - mu_edx)/(n()+l))
  
  predicted_ratings_b_movie_user <- 
    validation %>% 
    left_join(b_movie, by = "movieId") %>%
    left_join(b_user, by = "userId") %>%
    mutate(pred = mu_edx + b_movie + b_user) %>%
    .$pred
  
  return(RMSE(validation_New$rating,predicted_ratings_b_movie_user)) 
})

qplot(lambdas, rmses)  


lambda <- lambdas[which.min((rmses))]
lambda

```

the RMSE in this model is:

```{r edx11}
min(rmses)

rmse_results <- bind_rows(rmse_results,
                          data_frame(method="Regularized Movie and User Effect Model",RMSE = min(rmses) ))


```


##Results

We see the results with different model applicate and relative RMSE
```{r edx12}
rmse_results %>% knitr::kable()

```



## Conclusion
We see the best model to predict the rating movies is the fourth model where we applicate the regularization